Parallel air palletization and tour planning for simultaneous pickup and delivery

### 2022-11-20 latex/comp_or.tex
mpACO algorithm update

	\caption{ $SolveNode(mpACO, k, G, numProcs)$ } \label{alg:aco}
	\begin{algorithmic}[1]
		\State Let $G(V_k, E^Q_k)$
		\State $G^*(V_k, E^{N*}_k \cup E^Q_k) \gets Greedy(k, G, 1)$  \label{aco:greedy}
		\State $G^+(V_k, E^{N+}_k \cup E^Q_k) \gets G^*(V_k,E^{N*}_k \cup E^Q_k)$
		\State $A \gets$ a number of ants or parallel processes to be determined for each iteration \label{aco:ants}
		\State $ga^k_{ij}   \gets 0$, $1 \leq i \leq m$, $1 \leq j \leq n_k$  \label{aco:init1}
		\State $\tau^k_{ij} \gets 1$, $1 \leq i \leq m$, $1 \leq j \leq n_k$   \label{aco:init2}
		\State $\alpha    \gets 1.0$ \label{aco:alpha}
		\State $\beta     \gets 4.0$ \label{aco:beta}
		\State $\rho      \gets 0.2$ \label{aco:rho}
		\State $stagnant \gets 0$
		\While{($stagnant < 3$) {\bf and} ($runtime < 0.7s$)} \label{aco:while}

		\State $publish$
		\State $solutions \gets subscribe()$

			\For{$a \gets 1$ to $A$} \label{aco:g1}
				\State $G^a(V_k, E^{Na}_k \cup E^Q_k) \gets G^+(V_k, E^{N+}_k \cup E^Q_k)$
				\State $E \gets \{ (p_i,t^k_j) \mid (p_i,t^k_j) \not\in E^{Na}_k \}$
				\While{$|E| > 0$}
					\State $e \gets Roulette(E)$ biased through $ga^k_{ij}$  \label{aco:roulette}
					\State $E \gets E - \{e\}$		
					\If{$E^{Na}_k \cup \{e\}$ is feasible}
						\State $ E^{Na}_k \gets E^{Na}_k \cup \{e\} $ \label{aco:ant2}
					\EndIf	
				\EndWhile	
				\If{$f_s(G^{a}) > f_s(G^{+})$} \label{aco:gen1}
					\State $G^+(V_k, E^{N+}_k \cup E^Q_k) \gets G^a(V_k, E^{Na}_k \cup E^Q_k)$  \label{aco:gen2}
						\For{each $(p_i,t^k_j) \in E^{N+}_k$}
							\State $\tau^k_{ij} \gets \tau^k_{ij} - \Delta\tau^k_{ij}$ \label{aco:phero1}
							\State $ga^k_{ij} \gets (\tau^k_{ij})^\alpha + (\theta^k_{ij})^\beta$ \label{aco:ga1}
						\EndFor
				\EndIf 
			\EndFor \label{aco:g2}
			\If{$f_s(G^+) > f_s(G^*)$} \label{aco:best1}
				\State $stagnant \gets 0$
				\State $G^*(V_k, E^{N*}_k \cup E^Q_k) \gets G^+(V_k, E^{N+}_k \cup E^Q_k)$
				\For{each $(p_i,t^k_j) \in E^{N*}_k$}
					\State $\tau^k_{ij} \gets (1-\rho)\tau^k_{ij}$ \label{aco:rho1}
					\State $\tau^k_{ij} \gets \tau^k_{ij} + \Delta\tau^k_{ij}$ \label{aco:phero2}
					\State $ga^k_{ij} \gets (\tau^k_{ij})^\alpha + (\theta^k_{ij})^\beta$ \label{aco:ga2}
				\EndFor			
			\Else
				\State $stagnant \gets stagnant + 1$ \label{aco:best2}
			\EndIf
			\For{all $\tau^k_{ij}$} 
				\State $\tau^k_{ij} \gets (1-\rho)\tau^k_{ij}$ \label{aco:rho2}
			\EndFor	
		\EndWhile	
		\State ${\bf return}$ $G^*(V_k, E^{N*}_k \cup E^Q_k)$		
	\end{algorithmic}
\end{algorithm}

A greedy solution is obtained in line \ref{aco:greedy}, being initially stored in $G^*$\/ and $G^+$. The number $A$\/ of ants was empirically set to 4 (line \ref{aco:ants}), yielding solutions with a good relation quality/time. The $m \times n_k$\/ matrices $\tau^k_{ij}$ and $ga^k_{ij}$\/ (lines \ref{aco:init1}-\ref{aco:init2}) are used and updated by all ants.

In each generation, $A$\/ ants are created (lines \ref{aco:g1}-\ref{aco:g2}). The number of generations depends on the time limit or stagnation in improving results (line \ref{aco:while}). Each ant $a$\/ elaborates its allocation graph $G^a$, considering the current values of $ga^k_{ij}$. To enforce solution diversification, if $(p_i,t^k_j)$ is included in $G^a$\/ then pheromone $\tau^k_{ij}$\/ is decreased (line \ref{aco:phero1}). Similarly, the values of $ga^k_{ij}$\/ are also updated (line \ref{aco:ga1}), where the exponents $\alpha$\/ and $\beta$\/ were chosen empirically: $\alpha$ was set to 1.0 and $\beta$\/ to $4.0$\/ (lines \ref{aco:alpha}-\ref{aco:beta}). For further explanation, see \cite{DorigoManiezzoColorni1996}.

The best solution of each generation is stored in $G^+$\/ (lines \ref{aco:gen1}-\ref{aco:gen2}), and the update of $G^*$\/ and the stopping criterion are performed on lines \ref{aco:best1}-\ref{aco:best2}. Each edge $(p_i,t^k_j)$\/ in some solution has its corresponding pheromone updated in a cumulative way (line \ref{aco:phero2}), that is, more pheromone will be given to the most relevant edges, where the increment is defined in \ref{eq:delta_tau}. The same with $ga^k_{ij}$\/ (line \ref{aco:ga2}).

\begin{equation} \label{eq:delta_tau}
	\Delta\tau^k_{ij} = \frac{ ga^k_{ij}\times t^k_j.s }{ A \times t^k_j.v}
\end{equation}

Pheromone $\tau^k_{ij}$\/ evaporates (lines \ref{aco:rho1} and \ref{aco:rho2}), where $\rho$\/ is a evaporation rate, which can be seen as a mechanism that delays premature convergence towards a sub-optimal solution. $\rho$\/ was empirically set to $0.2$\/ (line \ref{aco:rho}).

>>>>

	\caption{ $SolveNode(mpACO, k, G, numProcs)$ } \label{alg:aco}
	\begin{algorithmic}[1]
		\State Let $G(V_k, E^Q_k)$
		\State $G^*(V_k, E^{N*}_k \cup E^Q_k) \gets Greedy(k, G, 1)$  \label{aco:greedy}
		\State $G^+(V_k, E^{N+}_k \cup E^Q_k) \gets G^*(V_k,E^{N*}_k \cup E^Q_k)$
		\State $A \gets numProcs$ \label{aco:numProcs}
		\State $ga^k_{ij}   \gets 0$, $1 \leq i \leq m$, $1 \leq j \leq n_k$  \label{aco:init1}
		\State $\tau^k_{ij} \gets 1$, $1 \leq i \leq m$, $1 \leq j \leq n_k$   \label{aco:init2}
		\State $\alpha    \gets 1.0$ \label{aco:alpha}
		\State $\beta     \gets 4.0$ \label{aco:beta}
		\State $\rho      \gets 0.2$ \label{aco:rho}
		\State $stagnant \gets 0$
		\While{ $stagnant < 3$ } \label{aco:while}

			\State $AntsQueue \gets \{\ \}$ \label{aco:AntsQueue}

			\For{$1$ {\bf to} $A$} \label{aco:AntsSolve}	
				\State $AntSolve(G^+,\ AntsQueue,\ ga^k_{ij})$
			\EndFor

			\State $counter \gets A$

			\While{$runtime < 0.7s$ {\bf and} $counter > 0$} \label{aco:while2}

				\State $G^{a} \gets AntsQueue.Pop()$	
				\If{$f_s(G^{a}) > f_s(G^{+})$}
					\State $G^+(V_k, E^{N+}_k \cup E^Q_k) \gets G^a(V_k, E^{Na}_k \cup E^Q_k)$ 
				\EndIf

				\State $counter \gets counter - 1$

			\EndWhile

			\If{$f_s(G^+) > f_s(G^*)$} \label{aco:best1}
				\State $stagnant \gets 0$
				\State $G^*(V_k, E^{N*}_k \cup E^Q_k) \gets G^+(V_k, E^{N+}_k \cup E^Q_k)$		
			\Else
				\State $stagnant \gets stagnant + 1$ 
			\EndIf
			\For{all $\tau^k_{ij}$} 
				\State $\tau^k_{ij} \gets (1-\rho)\tau^k_{ij}$ \label{aco:rho2}
				\State $\tau^k_{ij} \gets \tau^k_{ij} + \Delta\tau$ \label{aco:phero2}
				\State $ga^k_{ij} \gets (\tau^k_{ij})^\alpha + (\theta^k_{ij})^\beta$ \label{aco:ga2}				
			\EndFor	
		\EndWhile	
		\State ${\bf return}$ $G^*(V_k, E^{N*}_k \cup E^Q_k)$		
	\end{algorithmic}
\end{algorithm}

A greedy solution is obtained in line \ref{aco:greedy}, being initially stored in $G^*$\/ and $G^+$.
The number $A$\/ of ants was set to the number of parallel processes (line \ref{aco:numProcs}).

The $m \times n_k$\/ matrices $\tau^k_{ij}$ and $ga^k_{ij}$\/ (lines \ref{aco:init1}-\ref{aco:init2}) are used and updated by all ants. For further explanation, see \cite{DorigoManiezzoColorni1996}.

The number of iterations depends on the time limit or stagnation in improving results (line \ref{aco:while}).
At the beginning os each iteration (\ref{aco:AntsQueue}) a queue of ants ($AntsQueue$) is created to hold ants parallel solutions generated in the child processes.

The local solution $G^+$ is sent to A child processes (\ref{aco:AntsSolve}) almost simultaneously.
While there are any ant working on an solution or the run-time is still less then maximum (\ref{aco:while2}), it hangs in the $AntsQueue$ until there is a solution available.

If there is a global improvement (\ref{aco:best1}) the best so far solution ($G^*(V_k, E^{N*}_k \cup E^Q_k)$) is updated.
At the end of each iteration pheromone levels (\ref{aco:rho2}) are evaporated, than updated with the new calculated $\Delta\tau$ (line \ref{aco:phero2}), and according to Equation \ref{eq:delta_tau}, and the general attractiveness ( \ref{aco:ga2}) is also updated.

\begin{equation} \label{eq:delta_tau}
	\Delta\tau = \frac{f_s(G^{+}) - f_s(G^{*})}{A \times f_s(G^{*})}
\end{equation}

Many ACO applications update pheromone levels and the general attractiveness only when there is a solution improvement. Differently, this work updates in both cases, exploring also the negative learning approach proposed by \cite{math9040361}.

\begin{algorithm}[H] 
	\caption{ $AntSolve(G^+,\ AntsQueue,\ ga^k_{ij})$ } \label{alg:AntSolve}
	\begin{algorithmic}[1]
		\State $G^a(V_k, E^{Na}_k \cup E^Q_k) \gets G^+(V_k, E^{N+}_k \cup E^Q_k)$
		\State $E \gets \{ (p_i,t^k_j) \mid (p_i,t^k_j) \not\in E^{Na}_k \}$
		\While{$|E| > 0$}
			\State $e \gets Roulette(E)$ biased through $ga^k_{ij}$  \label{AntSolve:roulette}
			\State $E \gets E - \{e\}$		
			\If{$E^{Na}_k \cup \{e\}$ is feasible}
				\State $ E^{Na}_k \gets E^{Na}_k \cup \{e\} $ \label{AntSolve:ant}
			\EndIf	
		\EndWhile
		\State $AntsQueue \gets AntsQueue \cup G^a(V_k, E^{Na}_k \cup E^Q_k)$
	\end{algorithmic}
\end{algorithm}
